---
title: "GrpWk2_Econometrics"
author: "Group 3B"
date: "10/13/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We calculate the equilibrium FX for US Dollar/Naira for the monthly average exchange rate. The data is derived from <https://www.cbn.gov.ng/rates/exrate.asp>.

```{r}
# load packages and data
library(readxl)
library(vars)
library(timeSeries)
library(rugarch)
```

 
```{r}
# import the dataset
nairadolrates <- read_excel("MonthlyAverageExhRate13102020.xlsx", sheet = "MonthlyAverageExhRate13102020")
```

```{r}
acf(nairadolrates$`BDC (USD)`)
```
The above plot shows that the data is not stationary, hence, we make it stationary by making a first differencing i,e integrated of order 1.

```{r}
# find changes in interest rates by differencing:
diffnairausd <- diff(nairadolrates$`BDC (USD)`,trim=TRUE)
diffnairausd2 <- diff(nairadolrates$`IFEM (USD)`,trim=TRUE)

dnairausd = cbind(diffnairausd, diffnairausd2)
```


The autocorrelation function plot below shows a stationary differenced data.
```{r}
acf(diffnairausd)
```
```{r}
# simple correlations:
cor(dnairausd)
```
From the correlation table above, we have that the two rates are weakly correlated.
```{r}
# estimate a Vector Auto Regression
VAR_model <- VAR(dnairausd, lag.max=12, type = "none", ic = "AIC")
summary(VAR_model)
```
We try to use a impulse response function to interpret the table given above. Since the variables are quite many and we would like to observe the rule of parsimony.

```{r}
VAR_irf <- irf(VAR_model, n.ahead = 13,boot = TRUE, ci = 0.95)
plot(VAR_irf)
```
The above shows that the variables oscillates around the mean readily irrespective of whether a strong flunctuation is noticed or not.
```{r}
# lag order suggested by AIC

VAR_model$p
```
According to the above AIC, we have a lag order 9.


```{r}
# Cointegration Analysis
library(urca)
library(tsDyn)
```

```{r}
# store the various  monthly exchange rates:
BDC <- nairadolrates$`BDC (USD)`
IFEM <- nairadolrates$`IFEM (USD)`

rates = cbind(BDC, IFEM)
```

```{r}
# Johansen test
jotest1=ca.jo(rates, type="eigen", K=9, ecdet="none", spec="longrun")
summary(jotest1)
```
With the 10$\%$ test above, $r=1$
```{r warning=FALSE}
VECM_fit = VECM(rates,1,r = 1, include = "const",estim = "ML", LRinclude ="none")
summary(VECM_fit)
```
```









































